Bayesian Sparsification Methods for Deep Complex-valued Networks

### 2020-06-23 paper/paper.tex
fixed citations

% our contribution
Sparse Variational Dropout is a Bayesian Variational Inference method with automatic
parameter relevance determination effect. In this study we extend \emph{Sparse VD} to
$\cplx$VNN, inspired by the results of \citet{gale_state_2019}, and motivated by seldom
application of Bayesian Inference to $\cplx$-valued networks \citep{popa_complex-valued_2018}
and apparent scarcity of compression methods specific to them. We assess the performance-%
compression trade-off of the extension by conducting a large-scale numerical study on
image classification on MNIST-like and CIFAR10 datasets and music transcription on MusicNet.

@@

For an observed dataset $
  D = (x_i)_{i=1}^N
$ and statistical model $
  p(D \mid \omega)
    = \prod_i \log p(x_i \mid \omega)
    % $D$ conditional independence given parameters $\omega$
$ with parameters $\omega$ the Bayes rule transforms prior hypotheses $\pi(\omega)$
about the unknown distribution of model's parameters into the posterior distribution: $
  p(\omega \mid D) = \tfrac{p(D \mid \omega) \pi(\omega)}{p(D)}
$.
%
% In general the marginal likelihood of the dataset is intractable $
%   p(D) = \mathbb{E}_{\pi(\omega)} p(D \mid \omega)
% $.
Save for the relatively simple set-ups, either the posterior distribution itself or the
mathematical expectations it is involved in are analytically intractable or impractical
to compute numerically. Variational Inference (VI), proposed by \citet{jordan_introduction_1999},
can be used in such cases to make approximate inference. The approach finds an approximation
within some distribution family $q_\theta(\omega)$, which is closest to the true posterior
distribution in terms of Kullback-Leibler divergence: $
  KL(q_\theta(\omega) \| p(\omega \mid D))
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q_\theta}
      \log \tfrac{q_\theta(\omega)}{p(\omega \mid D)}
$. % The approximation family is chosen so that the expectations become tractable.
%
\citet{jordan_introduction_1999} show that this problem is equivalent to variational
\emph{maximization} of the \emph{Evidence Lower Bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \sum_{i=1}^N \mathbb{E}_{\omega \sim q_{\theta}}
        \log p(x_i \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the variational parameters $\theta$ and $\lambda$ parameterize the approximation
and the prior, respectively.
%
Kullback-Leibler and, by proxy, ELBO are standard objectives in VI, however it is possible
to use other objectives, provided the true posterior $p(\omega \mid D)$ is evaluated only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2018}.

@@

% The variance of the stochastic gradient directly affects convergence of SVI. 
Independently, \citet{kingma_auto-encoding_2014} proposed \emph{Stochastic Gradient
Variational Bayes}, which is an alternative efficient doubly stochastic estimator
applicable to models with parameters $\omega$ that are continuous random variables
amenable to the \emph{reparameterization trick}, i.e. $
  \omega \sim q_{\theta}(\omega)
$ is equivalent in distribution to $
  \omega = g_{\theta}(\varepsilon)
$ for some non-parametric random variable $
  \varepsilon \sim p(\varepsilon)
$ and $
  g(\varepsilon; \theta)
$ differentiable with respect to $\theta$. The estimator of \eqref{eq:elbo_general}
with $L$ reparameterized draws per element in the mini-batch of size $M$ is given by
\begin{equation}  \label{eq:sgvb_estimator}
  \widetilde{\mathcal{L}}(\theta; \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(x_i \mid \omega_i)b
      + \frac{N}{M L} \sum_{k,l}
        \log p(x_{i_k} \mid g(\varepsilon_{lk}; \theta))
      % + \frac{N}{M L} \sum_{k=1}^M \sum_{l=1}^L
      %   \log p(x_{i_k} \mid g(\varepsilon_{lk}; \theta))
        % \log p(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{x,\omega \sim B}
          % \log p(x \mid \omega)
    \,,
\end{equation}
where $(x_{i_k})_{k=1}^M$ is a random subsample from $D$ and $
  (\varepsilon_{lk})_{l=1}^L
$ are $k=1..M$ independent iid samples from $p_\varepsilon$. \citet{figurnov_implicit_2019}
extended the scope of the reparameterization gradients to include continuous distributions
such as Gamma and von Mises. To handle the case of non-reparameterizable $\omega$ in
doubly stochastic VI, e.g. discrete random parameters, \citet{titsias_local_2015} proposed
\emph{local expectation gradients}, which is a version of REINFORCE gradient estimator
\citep{williams_simple_1992} with variance reduced by careful use of dependence
structures in the model.
% (reparameterization trick, pathwise gradient) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
% rao-blackwellization (http://pyro.ai/examples/svi_part_iii.html)

@@

% on dropout
Dropout, proposed by \citet{hinton_improving_2012}, prevents overfitting by injecting
multiplicative binary noise into layer's weights, which breaks up co-adaptations that
could occur during training. \citet{wang_fast_2013} argued that the overall effect of
binary Dropout on the intermediate outputs can be approximated by a Gaussian with weight-input
dependent mean and variance via the Central Limit Theorem. \citet{srivastava_dropout:_2014}
proposed using independent $\mathcal{N}(1, 1)$ multiplicative noise, arguing that higher
entropy of a Gaussian has better regularizing effect. \citet{gal_dropout_2016} showed
that Dropout is a Bayesian approximation method with close ties to deep Gaussian Processes
that yields inexpensive model uncertainty estimates. In a study concerning multitask learning
\citet{cheung_superposition_2019} demonstrated the possibility of storing task-specific
parameters in non-destructive superposition within a single network. Regarding Dropout
their argument implies that if the single task setting is viewed as multitask learning
with replicated task, then by sampling uncorrelated binary masks Dropout acts as a
superposition method, utilizing the learning capacity of the network better.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex

>>>>

% our contribution
Sparse Variational Dropout is a Bayesian Variational Inference method with automatic
parameter relevance determination effect. In this study we extend \emph{Sparse VD} to
$\cplx$VNN, inspired by the results of \citet{gale_state_2019}, and motivated by seldom
application of Bayesian Inference to $\cplx$-valued networks \citep{popa_complex-valued_2017}
and apparent scarcity of compression methods specific to them. We assess the performance-%
compression trade-off of the extension by conducting a large-scale numerical study on
image classification on MNIST-like and CIFAR10 datasets and music transcription on MusicNet.

@@

For an observed dataset $
  D = (x_i)_{i=1}^N
$ and statistical model $
  p(D \mid \omega)
    = \prod_i \log p(x_i \mid \omega)
    % $D$ conditional independence given parameters $\omega$
$ with parameters $\omega$ the Bayes rule transforms prior hypotheses $\pi(\omega)$
about the unknown distribution of model's parameters into the posterior distribution: $
  p(\omega \mid D) = \tfrac{p(D \mid \omega) \pi(\omega)}{p(D)}
$.
%
% In general the marginal likelihood of the dataset is intractable $
%   p(D) = \mathbb{E}_{\pi(\omega)} p(D \mid \omega)
% $.
Save for the relatively simple set-ups, either the posterior distribution itself or the
mathematical expectations it is involved in are analytically intractable or impractical
to compute numerically. Variational Inference (VI), proposed by \citet{jordan_introduction_1999},
can be used in such cases to make approximate inference. The approach finds an approximation
within some distribution family $q_\theta(\omega)$, which is closest to the true posterior
distribution in terms of Kullback-Leibler divergence: $
  KL(q_\theta(\omega) \| p(\omega \mid D))
    % = \int \frac{dQ}{dP} \log{\frac{dQ}{dP}} P(d\omega)
    % = \int \frac{q(\omega)}{p(\omega)} \log{\frac{q(\omega)}{p(\omega)}} p(\omega) d\omega
    % = \mathbb{E}_{\omega \sim q}
    %   \log\frac{q(\omega)}{p(\omega)}
    = \mathbb{E}_{\omega \sim q_\theta}
      \log \tfrac{q_\theta(\omega)}{p(\omega \mid D)}
$. % The approximation family is chosen so that the expectations become tractable.
%
\citet{jordan_introduction_1999} show that this problem is equivalent to variational
\emph{maximization} of the \emph{Evidence Lower Bound} (ELBO)
\begin{equation}  \label{eq:elbo_general}
  \mathcal{L}(\theta; \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      + \sum_{i=1}^N \mathbb{E}_{\omega \sim q_{\theta}}
        \log p(x_i \mid \omega)
      % \mathbb{E}_{\omega \sim q_{\theta}}
        % \log\frac{q_\theta(\omega)}{\pi_{\lambda}(\omega)}
  % = \mathbb{E}_{\omega \sim q_{\theta}}
  %     \log{p_{\phi}(D \mid \omega) \pi_{\lambda}(\omega)}
  %   + \mathbb{H}(q_{\theta})
  \,,
\end{equation}
where the variational parameters $\theta$ and $\lambda$ parameterize the approximation
and the prior, respectively.
%
Kullback-Leibler and, by proxy, ELBO are standard objectives in VI, however it is possible
to use other objectives, provided the true posterior $p(\omega \mid D)$ is evaluated only
through $\log p(D \mid \omega)$ and $\log \pi(\omega)$, \citep{ranganath_operator_2016}.

@@

% The variance of the stochastic gradient directly affects convergence of SVI. 
Independently, \citet{kingma_auto-encoding_2014} proposed \emph{Stochastic Gradient
Variational Bayes}, which is an alternative efficient doubly stochastic estimator
applicable to models with parameters $\omega$ that are continuous random variables
amenable to the \emph{reparameterization trick}, i.e. $
  \omega \sim q_{\theta}(\omega)
$ is equivalent in distribution to $
  \omega = g_{\theta}(\varepsilon)
$ for some non-parametric random variable $
  \varepsilon \sim p(\varepsilon)
$ and $
  g(\varepsilon; \theta)
$ differentiable with respect to $\theta$. The estimator of \eqref{eq:elbo_general}
with $L$ reparameterized draws per element in the mini-batch of size $M$ is given by
\begin{equation}  \label{eq:sgvb_estimator}
  \widetilde{\mathcal{L}}(\theta; \lambda)
    = - KL(q_{\theta} \| \pi_{\lambda})
      % + \frac{N}{\lvert B \rvert} \sum_{i\in B}
        % \log p(x_i \mid \omega_i)b
      + \frac{N}{M L} \sum_{k,l}
        \log p(x_{i_k} \mid g(\varepsilon_{lk}; \theta))
      % + \frac{N}{M L} \sum_{k=1}^M \sum_{l=1}^L
      %   \log p(x_{i_k} \mid g(\varepsilon_{lk}; \theta))
        % \log p(D \mid \omega_{l})
        %   \Big\vert_{\omega_{l} = g(\varepsilon_{l}; \theta)}
      % =
      % - KL(q_{\theta} \| \pi)
      % + N \mathbb{E}_{B \sim \{D\}_M \otimes q_\theta^M}
        % \hat{\mathbb{E}}_{x,\omega \sim B}
          % \log p(x \mid \omega)
    \,,
\end{equation}
where $(x_{i_k})_{k=1}^M$ is a random subsample from $D$ and $
  (\varepsilon_{lk})_{l=1}^L
$ are $k=1..M$ independent iid samples from $p_\varepsilon$. \citet{figurnov_implicit_2018}
extended the scope of the reparameterization gradients to include continuous distributions
such as Gamma and von Mises. To handle the case of non-reparameterizable $\omega$ in
doubly stochastic VI, e.g. discrete random parameters, \citet{titsias_local_2015} proposed
\emph{local expectation gradients}, which is a version of REINFORCE gradient estimator
\citep{williams_simple_1992} with variance reduced by careful use of dependence
structures in the model.
% (reparameterization trick, pathwise gradient) $q_{\theta}(d\omega)$ is a push-forward
% of $p(d\varepsilon)$ by a differentiable map $g(\varepsilon; \theta)$.
% % http://stillbreeze.github.io/REINFORCE-vs-Reparameterization-trick/#fn:1
% rao-blackwellization (http://pyro.ai/examples/svi_part_iii.html)

@@

% on dropout
Dropout, proposed by \citet{hinton_improving_2012}, prevents overfitting by injecting
multiplicative binary noise into layer's weights, which breaks up co-adaptations that
could occur during training. \citet{wang_fast_2013} argued that the overall effect of
binary Dropout on the intermediate outputs can be approximated by a Gaussian with weight-input
dependent mean and variance via the Central Limit Theorem. \citet{srivastava_dropout_2014}
proposed using independent $\mathcal{N}(1, 1)$ multiplicative noise, arguing that higher
entropy of a Gaussian has better regularizing effect. \citet{gal_dropout_2016} showed
that Dropout is a Bayesian approximation method with close ties to deep Gaussian Processes
that yields inexpensive model uncertainty estimates. In a study concerning multitask learning
\citet{cheung_superposition_2019} demonstrated the possibility of storing task-specific
parameters in non-destructive superposition within a single network. Regarding Dropout
their argument implies that if the single task setting is viewed as multitask learning
with replicated task, then by sampling uncorrelated binary masks Dropout acts as a
superposition method, utilizing the learning capacity of the network better.
% many identical copies of the same task; \cite[Appendix A.1]{cheung_superposition_2019}
% (fast dropout gauss-approx) wang-manning.tex
% (gauss dropout) srivastava-et-al.tex
